---
title: "Universality of Uniform Distribution"
author: "Rajveer"
date: "2025-09-28"
output: github_document
---

```{r setup, include=FALSE}
# Set global chunk options
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
set.seed(123)  # for reproducibility
```

## 1. Simulate Uniform Random Numbers

We first generate uniform random numbers on $[0,1]$.

```{r uniform-simulation}
# Number of samples
n <- 10000

# Generate uniform random numbers on [0,1]
U <- runif(n)
head(U)
```

---

## 2. Transform Uniform to Other Distributions

### a) Exponential Distribution

If $X \sim \text{Exponential}(\lambda)$, the CDF is:

$$
F(x) = 1 - e^{-\lambda x}, \quad x \ge 0
$$

The inverse CDF is:

$$
F^{-1}(u) = -\frac{\log(1-u)}{\lambda}
$$

We can transform uniform samples to exponential samples:

```{r exponential-transformation}
lambda <- 2
X_exp <- -log(1 - U) / lambda
head(X_exp)
```

---

### b) Normal Distribution

For a standard normal distribution $X \sim N(0,1)$, we can use the built-in `qnorm()` function (inverse CDF):

```{r normal-transformation}
X_norm <- qnorm(U)
head(X_norm)
```

---

### c) Any Other Distribution

For any distribution with CDF $F$, apply its inverse CDF.

**Example: Uniform distribution on \[a,b]:**

```{r uniform-a-b}
a <- 5
b <- 10
X_uniform <- a + (b - a) * U
head(X_uniform)
```

---

## 3. Visualize the Distributions

```{r plots, fig.width=8, fig.height=6}
par(mfrow=c(2,2))
hist(U, breaks=30, main="Uniform [0,1]", col="lightblue", xlab="U")
hist(X_exp, breaks=30, main="Exponential", col="lightgreen", xlab="X_exp")
hist(X_norm, breaks=30, main="Normal", col="lightpink", xlab="X_norm")
hist(X_uniform, breaks=30, main="Uniform [5,10]", col="lightyellow", xlab="X_uniform")
```

---

This demonstrates how **one set of uniform random numbers can be transformed into many different distributions** using the **inverse CDF method**.
